---
title: "Drink and Thrive? The economics of binge-drinking in Germany"
subtitle: "CSSR Presentation"
author: "Torben and Alex"
date: "02 December 2016"
output: ioslides_presentation
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Clear Global environment
rm(list=ls())

# Setting Working directory
try(setwd("/home/torben/GIT/Pair_Assignment_2"), silent = TRUE)
try(setwd("D:/Eigene Datein/Dokumente/Uni/Hertie/Materials/Collaborative Social Science Data Analysis/CSSR_Project"), silent = TRUE)

source("main.R")

# Collect packages/libraries we need:
packages <- c("fontcm", "psych")
# can we add a short list so we have an overview for what we need the packages?
# psych allows for summary statistics

# install packages if not installed before
for (p in packages) {
  if (p %in% installed.packages()[,1]) {
    require(p, character.only=T)
  }
  else {
    install.packages(p, repos="http://cran.rstudio.com", dependencies = TRUE)
    require(p, character.only=T)
  }
}
rm(p, packages)

# Initiate Label Names
statelist_code <- c("DE-BW", "DE-BY", "DE-BE", "DE-BB", "DE-HB", "DE-HH", 
                    "DE-HE", "DE-MV", "DE-NI", "DE-NW", "DE-RP", "DE-SL",
                    "DE-SN", "DE-ST", "DE-SH", "DE-TH")

statelist_name <- c("Baden-Württemberg", "Bayern", "Berlin", "Brandenburg", 
                    "Bremen", "Hamburg", "Hessen", "Mecklenburg-Vorpommern", 
                    "Niedersachsen", "Nordrhein-Westfalen", "Rheinland-Pfalz", 
                    "Saarland", "Sachsen", "Sachsen-Anhalt", 
                    "Schleswig-Holstein", "Thüringen")
statelist.code.short <- substr(statelist_code, 4, 5)
```

## What / Structure

Here we go!!!!

## Background

- Bullet 1
- Bullet 2
- Bullet 3

## Data

```{r cars, echo = TRUE}
summary(cars)
```

## Empiricial Strategy

- Dependent variable: Alcohol-related hospitalization (pp 1000)

**Question 1: Differences between states**

Method: Multiple regression

- *Variation*: Levels and differences
- *Robustness check:* Different years and long vs. short-term alcohol diagnosis

**Question 2: Effect of night-sale-ban**

Method: Difference-in-Difference

- *Variation*: Control variables and number of periods
- *Robustness check:* different control group compositions

## Data Visuals - F10.0 Cases in Germany

```{r echo = FALSE, fig.width=4, fig.height=5, out.extra='style="float:left"'} 
spplot(map.Germany.2000.F100_p1000, 
       zcol = "value", 
       col.regions = colors,
       main=list(label="2000: F10.0 diagnoses\nper 1000 pp all age groups"),
       at = seq(0, 2.2, by = 0.1))
```
```{r echo = FALSE, fig.width=4, fig.height=5, out.extra='style="float:right"'} 
spplot(map.Germany.2014.F100_p1000, 
       zcol = "value", 
       col.regions = colors,
       main=list(label="2014: F10.0 diagnoses\nper 1000 pp all age groups"),
       at = seq(0, 2.2, by = 0.1),
       colorkey=FALSE)
```

## Data Visuals -  F10.0 per Age Group

```{r echo = FALSE, fig.width=8, fig.height=5} 
# Bar diagram: F10.0 cases per 1000 per age group in DE-DE
PL1 <- TOTAL %>% filter(GENDER == "all", AGE!="all", STATE=="DE-DE")
ggplot(data=PL1, aes(x=AGE, y=F100_p1000)) +
  geom_bar(stat="identity", fill = "#1E90FF") +
  xlab("Age Group") +
  ylab("F10.0 cases per 1000 people") +
  ggtitle("F10.0 cases in Germany \n per 1000 per age group") +
  guides(fill=FALSE) +
  theme(axis.text.x  = element_text(angle=90, vjust=0.5),
        axis.title.y=element_text(margin=margin(10,10,0,10)))
```

## Data Visuals - K70 per Age Group
```{r echo = FALSE, fig.width=8, fig.height=5} 
# Bar diagram: K70 cases per 1000 per age group
ggplot(data=PL1, aes(x = AGE, y = K70_p1000, fill = AGE)) +
  geom_bar(stat="identity", fill = "#1E90FF") +
  xlab("Age Group") +
  ylab("K70 cases per 1000 people") +
  ggtitle("K70 cases in Germany \n per 1000 per age group") +
  guides(fill=FALSE) +
  theme(axis.text.x  = element_text(angle=90, vjust=0.5),
        axis.title.y=element_text(margin=margin(10,10,0,10)))
```


## Results

## Conclusion

## Appendix 1

```{r Model1, results = "asis", echo = FALSE}
mod1.labels <- c("GDP per capita", "Unemployment rate", "Beer tax", "Population density", "(Intercept)")
stargazer::stargazer(mod1.2000, mod1.2007, mod1.2014,
                     covariate.labels = mod1.labels,
                     column.labels = c("2000", "2007", "2014"),
                     model.numbers  = FALSE,
                     dep.var.labels = "F10.0 Diagnoses per 1000 capita",
                     title = 'Regression results for Model 1 for 2000, 2007 and 2014',
                     digits = 2, type = 'html', header = FALSE)
```

## Appendix 2

```{r Model2, results = "asis", echo = FALSE} 
stargazer::stargazer(mod2.F100, mod2.F102, mod2.K70,
                     column.labels = c("F10.0", "F10.2", "K70"),
                     covariate.labels = c("GDP Change", "Unemployment Change", "Beer Tax Change", "(Intercept)"),
                     model.numbers  = FALSE,
                     dep.var.labels = c("Change in F10.0","Change in F10.2","Change in K70"),
                     title = 'Regression results for Model 2 with first differenced data',
                     digits = 2, type = 'html', header = FALSE)
```

## Appendix 3

```{r Model3, results = "asis", echo = FALSE} 
stargazer::stargazer(mod3.16, mod3.13, 
                     column.labels = c("All States", "All but City States"),
                     covariate.labels = c("DE-BW dummy", "Post-treatment dummy", "Interaction", "(Intercept)"),
                     model.numbers  = FALSE,
                     dep.var.labels = c("F10.0 cases per 1000 PP"),                    
                     title = 'Model 3: simple differences-in-differences',
                     digits = 2, type = 'html', header = FALSE)

```

## Appendix 4

```{r Model4, results = "asis", echo = FALSE} 
stargazer::stargazer(mod4, 
                     column.labels = c("All States"),
                     covariate.labels = c("DE-BW dummy", "Post-sales ban dummy", 
                                          "GDP per capita", "Youth uneployment rate", 
                                          "Interaction dummy", "(Intercept)"),
                     model.numbers  = FALSE,
                     dep.var.labels = c("F10.0 cases per 1000 PP"),                           
                     title = 'Model 4: simple differences-in-differences with controls',
                     digits = 2, type = 'html', header = FALSE)
```

## Appendix 5

```{r Model5, results = "asis", echo = FALSE} 

stargazer::stargazer(mod5, 
                     column.labels = c("All States"),
                     covariate.labels = c("DE-BW ban dummy", "Youth uneployment rate", "GDP per capita"),
                     model.numbers  = FALSE,
                     dep.var.labels = c("F10.0 cases per 1000 PP"),                           
                     title = 'Model 5: multi-period panel differences-in-differences',
                     digits = 2, type = 'html', header = FALSE)
```