---
title: "Drink and Thrive? The economics of binge-drinking in Germany"
subtitle: "CSSR Presentation"
author: "Torben and Alex"
date: "02 December 2016"
output: ioslides_presentation
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Clear Global environment
rm(list=ls())

# Setting Working directory
try(setwd("/home/torben/GIT/Pair_Assignment_2"), silent = TRUE)
try(setwd("D:/Eigene Datein/Dokumente/Uni/Hertie/Materials/Collaborative Social Science Data Analysis/CSSR_Project"), silent = TRUE)

source("main.R")

# Collect packages/libraries we need:
packages <- c("fontcm", "psych")
# can we add a short list so we have an overview for what we need the packages?
# psych allows for summary statistics

# install packages if not installed before
for (p in packages) {
  if (p %in% installed.packages()[,1]) {
    require(p, character.only=T)
  }
  else {
    install.packages(p, repos="http://cran.rstudio.com", dependencies = TRUE)
    require(p, character.only=T)
  }
}
rm(p, packages)

# Initiate Label Names
statelist_code <- c("DE-BW", "DE-BY", "DE-BE", "DE-BB", "DE-HB", "DE-HH", 
                    "DE-HE", "DE-MV", "DE-NI", "DE-NW", "DE-RP", "DE-SL",
                    "DE-SN", "DE-ST", "DE-SH", "DE-TH")

statelist_name <- c("Baden-Württemberg", "Bayern", "Berlin", "Brandenburg", 
                    "Bremen", "Hamburg", "Hessen", "Mecklenburg-Vorpommern", 
                    "Niedersachsen", "Nordrhein-Westfalen", "Rheinland-Pfalz", 
                    "Saarland", "Sachsen", "Sachsen-Anhalt", 
                    "Schleswig-Holstein", "Thüringen")
statelist.code.short <- substr(statelist_code, 4, 5)
```

## What / Structure

Here we go!!!!

## Background

- Bullet 1
- Bullet 2
- Bullet 3

## Data

```{r cars, echo = TRUE}
summary(cars)
```

## Empiricial Strategy

- Dependent variable: Alcohol-related hospitalization (pp 1000)

**Question 1: Differences between states**

Method: Multiple regression

- *Variation*: Levels and differences
- *Robustness check:* Different years and long vs. short-term alcohol diagnosis

**Question 2: Effect of night-sale-ban**

Method: Difference-in-Difference

- *Variation*: Control variables and number of periods
- *Robustness check:* different control group compositions


## Data Visuals

```{r echo = FALSE, out.width=c('300px', '300px'), fig.show='hold'} 
spplot(map.Germany.2000.F100_p1000, 
       zcol = "value", 
       col.regions = colors,
       main=list(label="2000: F10.0 diagnoses\nper 1000 pp all age groups"),
       at = seq(0, 2.2, by = 0.1))
spplot(map.Germany.2014.F100_p1000, 
       zcol = "value", 
       col.regions = colors,
       main=list(label="2014: F10.0 diagnoses\nper 1000 pp all age groups"),
       at = seq(0, 2.2, by = 0.1),
       colorkey=FALSE)
```

## Results

## Conclusion